Market-based Mechanisms for Resource Scheduling

### 2013-12-18 paper/paper.tex
Made table

We will use the suggestive term “realistic conditions” to mean a situation in which any of the following occur (i.e., our negative results apply when any of these occur, and the setting is otherwise the simplest setting of static, homogeneous demands for homogeneous resources):

@@

We say that a system is “demand-based” if it elicits a demand from each user and makes assignments on the basis of the stated demands. For example, DRF, asset allocation, and CEEI all fall into this class.

@@

Suppose that there are two users and that at each time each user has enough tasks to consume all of the clusters’ resources. Tasks arrive at multiples of time T and last for T rounds. Moreover, suppose that during each interval of time T each user has an independently and uniformly randomly generated priority which is either “high” or “low.” The only characteristic of these priorities is that, all else equal, they would prefer run one extra high priority task at the expense of one fewer low priority task.

Since the scheduler is demand-based and the user’s priority levels are changing across time, the allocation must be independent of the priority of the users. So if each user receives a non-zero share of the resources, then each user must receive a non-zero share of the resources while they are in the “high” priority state and the other user is in the “low” priority state.

@@

DRF satisfies a weaker form of strategy-proofness which we call ``local strategy-proofness,'' 
such that a user is incentivized to honestly state their demands given their beliefs about the effect of those demands on the scheduler’s future behavior. 
This differs from the ordinary notion in that a user may need to reason strategically about future rounds of allocation in order to determine
their demand, rather than honestly reporting their real demand.
%XXX replace user by framework
%XXX fold this into the clearer general discussion of loca guarantees

@@

\begin{itemize}
\item \textbf{Strategy-proofness:} consider the setting where a framework’s top priority is achieving some particularly valuable resource X 
which is currently in use, though they have other tasks which they value less. To be most extreme, suppose the user values any outcome in which they receive X 
higher than any outcome in which they do not. 
A user can arrange to receive X by strategically passing up resource offers which they could use, in order to maximize 
their probability of being offered the resource X when it first becomes free. 
\item \textbf{Pareto-efficiency:} This is a corollary of theorem~\ref{demand-bad}. 
In the case of inhomogeneous resources, suppose that two users have streams of tasks, some of which run faster on resources of type A and some of which run faster on resources of type B. 
Then no demand-based scheduler can ensure that tasks are always placed on the resources on which they run fastest.
\item \textbf{Envy-freeness and compatibility with sharing incentive:} Suppose that one user consumes $\frac 13$ of the resources in the cluster and another consumes $\frac 12$ of the resources, 
but a particular $frac 1{10}$ of the resources in the cluster are $10$ times more valuable (to every user) than they appear. 
This could be based on hardware characteristics to which the scheduler is oblivious. 
The user who consumes $\frac 13$ of the resources receives all resource offers first, and so claims all of the especially valuable resources. 
But this means that this user receives significantly more total value than the other user, violating both envy-freeness and the sharing incentive. 
\end{itemize}

The results above hold only for demand-based policies.
The following result shows that the strong guarantees of the sharing
incentive and Pareto efficiency are actually incompatible in general.
This motivates us to introduce weaker notions in the next section.
\begin{theorem}\label{no-efficiency}
%XXX change the appearance of "theorem 3" below to reference this one
The sharing incentive and Pareto efficiency are incompatible under realistic conditions.
\end{theorem}
\begin{pproof}
Consider a setting in which there are two phases and two users. 
The second phase differs from the first in one of two ways: 
either it is the case that user 2 cares more about their tasks in phase 2 than about their tasks in phase 1, 
or it is the case that user 1 cares more about their tasks in phase 2 than about their tasks in phase 1. 
By “cares more about” we mean simply that a user would, all else equal, 
prefer to complete those tasks more rapidly at the expense of completing other tasks less rapidly.

@@

We define a user’s fair share of the cluster to be a proportional share of each resource, with the proportion determined by the ratio between the total amount paid by that user, and the amount paid by all users over the cluster.

@@

%XXX link this into the general discussion
As with the envy-free analysis of DRF, we need to assume that either there are no complementarities between resources, or that the user’s demand for a resource takes into account their expectations about future allocations (e.g. if resources R and Q are only valuable together, we either assume that the scheduler packages them together into a single resource or that the frameworks are able to evaluate the value of r based on the probability that they will be able to receive Q).

@@

\begin{pproof}
We will show by induction that for any number of rounds, user A would prefer to have received user A’s allocation at user A’s price than user B’s allocation at user B’s price. Consider a particular round, after which user A receives some resources $R_A$ at a price $p_A$ and user B receives resources $R_B$ at a price $p_B$.

By the definition of our auction, the price $p_B$ exceeds user A’s willingness to pay for resources $R_B$ (or else user A’s proxy would have continued bidding and the auction would not have concluded), while the price $p_A$ is less than user A’s willingness to pay for resources $R_A$ (or else user A’s proxy would have stopped bidding earlier). In particular, user A prefers their package to user B’s package. Since at the start of the round user A preferred their allocation at their price to user B’s allocation at user B’s price, this property also holds at the end of the round. In this step we rely on our assumption that user A’s valuations for their own holdings do not change after the auction is announced (otherwise it may be possible for user A to become envious in between rounds as more information is revealed, rather than during a round of the auction). We note in passing that similar assumptions are required to prove that DRF is envy-free in the dynamic setting. []
\end{pproof}

@@

\begin{theorem}[Local strategy-proofness] 
Assuming that users are indifferent to changes in each others’ budgets,
and that users’ willingness to pay 
incorporates accurate expectations of effects on future allocations, 
the optimal behavior in our auction is to honestly declare willingness to pay.
\end{theorem}

@@

The defining characteristic of our auction is that $p_i$ is equal to the loss in other participants’ welfare which results from awarding user A the resource bundle $R_i$. 
WLOG suppose that $p_2> p_1$.
This means that the auctioneer would need to reduce the prices charged to the other users by at most $(p_2 - p_1)$ in order to change user A’s bundle from $R_1$ to $R_2$. 
But by hypothesis, user A would be willing to pay more than $(p_2 - p_1)$ in order to make this change. 
Since $B_1$ was an honest bid, the auctioneer can raise A’s price by more than $(p_2 - p_1)$, 
and since the auctioneer continues to make favorable modifications until no more are available, it will do so.

@@

%XXX make notes about approximations, say we are analyzing the intractable optimization and that statements hold with appropriate caveats, e.g. you aren’t Pareto efficient if you can’t do the optimization exactly, but you are as Pareto efficient as you can realistically be. Since optimization is computationally hard, the only way we can avoid this caveat is by ignoring it.]
%XXX make notes about the notion of ``local'' guarantees

@@

\begin{table*}[t!]
\begin{center}
    \begin{tabular}{| l | l | l | l | l | l |}
    \hline
    Desired & \small DRF & \small DRF & \small Market-based  & something & something\\
    Properties & \small(simple) & \small (complex) & \small Mechanism  & something & something\\
    \hline
    \small Envy-Freeness & \cmark & \cmark  & \cmark & \cmark   & \cmark\\
    \small Pareto-Efficiency & 5 & 6  & 4  & \cmark  & \cmark\\
    \small Resource Monotonicity & 8 & 9  & 4  & \cmark  & \cmark \\
    7 & 8 & 9  & 4  & \cmark  & \cmark \\
    7 & 8 & 9  & 4  & \cmark  & \cmark\\
    7 & 8 & 9  & 4 & \cmark  & \cmark\\
    \hline
    \end{tabular}
\end{center}
  \caption{A simple table}
  \label{table:props}
\end{table*}

>>>>

We will use the suggestive term ``realistic conditions'' to mean a situation in which any of the following occur (i.e., our negative results apply when any of these occur, and the setting is otherwise the simplest setting of static, homogeneous demands for homogeneous resources):

@@

We say that a system is ``demand-based'' if it elicits a demand from each user and makes assignments on the basis of the stated demands. For example, DRF, asset allocation, and CEEI all fall into this class.

@@

Suppose that there are two users and that at each time each user has enough tasks to consume all of the clusters' resources. Tasks arrive at multiples of time T and last for T rounds. Moreover, suppose that during each interval of time T each user has an independently and uniformly randomly generated priority which is either ``high'' or ``low.'' The only characteristic of these priorities is that, all else equal, they would prefer run one extra high priority task at the expense of one fewer low priority task.

Since the scheduler is demand-based and the user's priority levels are changing across time, the allocation must be independent of the priority of the users. So if each user receives a non-zero share of the resources, then each user must receive a non-zero share of the resources while they are in the ``high'' priority state and the other user is in the ``low'' priority state.

@@

DRF satisfies a weaker form of strategy-proofness which we call ``local strategy-proofness,'' 
such that a user is incentivized to honestly state their demands given their beliefs about the effect of those demands on the scheduler's future behavior. 
This differs from the ordinary notion in that a user may need to reason strategically about future rounds of allocation in order to determine
their demand, rather than honestly reporting their real demand.
%XXX replace user by framework
%XXX fold this into the clearer general discussion of loca guarantees

@@

\begin{itemize}
\item \textbf{Strategy-proofness:} consider the setting where a framework's top priority is achieving some particularly valuable resource X 
which is currently in use, though they have other tasks which they value less. To be most extreme, suppose the user values any outcome in which they receive X 
higher than any outcome in which they do not. 
A user can arrange to receive X by strategically passing up resource offers which they could use, in order to maximize 
their probability of being offered the resource X when it first becomes free. 
\item \textbf{Pareto-efficiency:} This is a corollary of theorem~\ref{demand-bad}. 
In the case of inhomogeneous resources, suppose that two users have streams of tasks, some of which run faster on resources of type A and some of which run faster on resources of type B. 
Then no demand-based scheduler can ensure that tasks are always placed on the resources on which they run fastest.
\item \textbf{Envy-freeness and compatibility with sharing incentive:} Suppose that one user consumes $\frac 13$ of the resources in the cluster and another consumes $\frac 12$ of the resources, 
but a particular $frac 1{10}$ of the resources in the cluster are $10$ times more valuable (to every user) than they appear. 
This could be based on hardware characteristics to which the scheduler is oblivious. 
The user who consumes $\frac 13$ of the resources receives all resource offers first, and so claims all of the especially valuable resources. 
But this means that this user receives significantly more total value than the other user, violating both envy-freeness and the sharing incentive. 
\end{itemize}

The results above hold only for demand-based policies.
The following result shows that the strong guarantees of the sharing
incentive and Pareto efficiency are actually incompatible in general.
This motivates us to introduce weaker notions in the next section.
\begin{theorem}\label{no-efficiency}
%XXX change the appearance of "theorem 3" below to reference this one
The sharing incentive and Pareto efficiency are incompatible under realistic conditions.
\end{theorem}
\begin{pproof}
Consider a setting in which there are two phases and two users. 
The second phase differs from the first in one of two ways: 
either it is the case that user 2 cares more about their tasks in phase 2 than about their tasks in phase 1, 
or it is the case that user 1 cares more about their tasks in phase 2 than about their tasks in phase 1. 
By ``cares more about'' we mean simply that a user would, all else equal, 
prefer to complete those tasks more rapidly at the expense of completing other tasks less rapidly.

@@

We define a user's fair share of the cluster to be a proportional share of each resource, with the proportion determined by the ratio between the total amount paid by that user, and the amount paid by all users over the cluster.

@@

%XXX link this into the general discussion
As with the envy-free analysis of DRF, we need to assume that either there are no complementarities between resources, or that the user's demand for a resource takes into account their expectations about future allocations (e.g. if resources R and Q are only valuable together, we either assume that the scheduler packages them together into a single resource or that the frameworks are able to evaluate the value of r based on the probability that they will be able to receive Q).

@@

\begin{pproof}
We will show by induction that for any number of rounds, user A would prefer to have received user A's allocation at user A's price than user B's allocation at user B's price. Consider a particular round, after which user A receives some resources $R_A$ at a price $p_A$ and user B receives resources $R_B$ at a price $p_B$.

By the definition of our auction, the price $p_B$ exceeds user A's willingness to pay for resources $R_B$ (or else user A's proxy would have continued bidding and the auction would not have concluded), while the price $p_A$ is less than user A's willingness to pay for resources $R_A$ (or else user A's proxy would have stopped bidding earlier). In particular, user A prefers their package to user B's package. Since at the start of the round user A preferred their allocation at their price to user B's allocation at user B's price, this property also holds at the end of the round. In this step we rely on our assumption that user A's valuations for their own holdings do not change after the auction is announced (otherwise it may be possible for user A to become envious in between rounds as more information is revealed, rather than during a round of the auction). We note in passing that similar assumptions are required to prove that DRF is envy-free in the dynamic setting. []
\end{pproof}

@@

\begin{theorem}[Local strategy-proofness] 
Assuming that users are indifferent to changes in each others' budgets,
and that users' willingness to pay 
incorporates accurate expectations of effects on future allocations, 
the optimal behavior in our auction is to honestly declare willingness to pay.
\end{theorem}

@@

The defining characteristic of our auction is that $p_i$ is equal to the loss in other participants' welfare which results from awarding user A the resource bundle $R_i$. 
WLOG suppose that $p_2> p_1$.
This means that the auctioneer would need to reduce the prices charged to the other users by at most $(p_2 - p_1)$ in order to change user A's bundle from $R_1$ to $R_2$. 
But by hypothesis, user A would be willing to pay more than $(p_2 - p_1)$ in order to make this change. 
Since $B_1$ was an honest bid, the auctioneer can raise A's price by more than $(p_2 - p_1)$, 
and since the auctioneer continues to make favorable modifications until no more are available, it will do so.

@@

%XXX make notes about approximations, say we are analyzing the intractable optimization and that statements hold with appropriate caveats, e.g. you aren't Pareto efficient if you can't do the optimization exactly, but you are as Pareto efficient as you can realistically be. Since optimization is computationally hard, the only way we can avoid this caveat is by ignoring it.]
%XXX make notes about the notion of ``local'' guarantees

@@

\begin{table*}[t!]
\begin{center}
    \begin{tabular}{| l | l | l | l | l | l |}
    \hline
    Desired & \small DRF & \small DRF & \small Market-based  \\
    Properties & \small(simple) & \small (realistic) & \small Mechanism \\
    \hline
    \small Envy-Free & \cmark &  & \cmark \\
    \small Pareto-Efficient & \cmark & Locally & Locally  \\
    \small Strategy-proof & \cmark & Locally  & Locally \\
    \small Sharing incentive & \cmark &  & \cmark\\
    \hline
    \end{tabular}
\end{center}
  \caption{A simple table}
  \label{table:props}
\end{table*}
